---
title: "Markdown file"
author: "Jenny Ly"
date: "25/09/2022"
output: html_document
---

```{r}
library(here)
# Import data
hyper <- read.csv(here("data","hyper.csv"))

#Package used 
library(VIM)

```

```{r}
#Read (Statistical Analysis data)
dim(hyper)
##1000 rows, 5 columns
```

```{r}
head(hyper)
str(hyper)
## no character variables, all integer variables ==> errors in data
## gender:  categorical variable
## hypertension: categorical variable
## bmi: categorical variable
## missing values (NA) appear

summary(hyper)
## min in "age_mn", "hypertension" , "bmi" = 0 
## missing values and incorrect variable's definition affects summary's result 

```

```{r}
# Check missing data
aggr(hyper,sortVar=TRUE)$combination
##  missing data:  "age_mn" - 16,  "hypertension"-271 , "bmi"-166
## proportion of missing data: "age_mn" 1.6% "hypertension" 27.1% "bm" 16.6%

## Potential errors in data include missing value in data and undefined variables
```

```{r}
# Define categorical variables 
hyper$gender <- factor(hyper$gender)
table(hyper$gender) #2 levels: 1,2

hyper$hypertension <- factor(hyper$hypertension)
table(hyper$hypertension) # 2 levels: 0,1

hyper$bmi <-factor(hyper$bmi)
table(hyper$bmi) # 4 levels: 0,1,2,3
```

```{r}
# statistical analysis for new data set
new_hyper <- na.omit(hyper) #new data set without missing values
```

```{r}
#Age: 
summary(new_hyper$age_mn)
### median age without missing values: 328.0, range from 98-1012
IQR(new_hyper$age_mn)
###interquartile range of age: 440
```

### Normal Distribution

```{r}
#Gender
table(new_hyper$gender)
table(hyper$gender)
## Even there is no missing values found in gender, but there are 107 rows 
##which contain missing values are delleted from original data set.
##Therefore number of male and female is deducted 
gender_prop <- 100*table(new_hyper$gender)/length(new_hyper$gender)
## 50.6% of male , 49.4% of female 
```

```{r}
#hypertension
table(new_hyper$hypertension)
hyper_prop <- 100*table(new_hyper$hypertension)/length(new_hyper$hypertension)
###86.3%  no hypertension, 13.7% with hypertension

```


```{r}
#bmi
table(new_hyper$bmi)
bmi_prop <- 100*table(new_hyper$bmi)/length(new_hyper$bmi)
### 20.3% obese, 30% overweight, 39.1% normal, 10.5% underweight

```

### Age Plot

```{r,echo=FALSE}
#*Plot*
##Age
par(mfrow = c(2,2))
boxplot(new_hyper$age_mn, main = "age by month")
hist(new_hyper$age_mn,main = "age",
     xlab= "age by month")
hist(new_hyper$age_mn,probability = 1, main = "age",
     xlab= "age by month")
lines(density(new_hyper$age_mn))

```

no outliners showed in box plot graph, Min:96, max:1012, median: 328 the graph indicates right skew, transformation for data is considered 

### Proportion graph

```{r,echo=FALSE}
##Hypertension 
par(mfrow = c(1, 2))
barplot(table(new_hyper$hypertension),main = "Hypertension",
        names.arg = c("no-hypertension","hypertension"))
##Bmi
barplot(table(new_hyper$bmi),main = "bmi",
        names.arg = c("obese","overweigth","normal","underweigth"))
```

### Relationship between graphs

```{r,echo=FALSE}
par(mfrow = c(2, 2))
##Relationship between bmi & hypertension
compare <- table(new_hyper$bmi,new_hyper$hypertension)
mosaicplot(compare,color = c("light yellow","purple"),
           main = "Hypertension vs Bmi")
###high percentage of people with hypertension without obesity, 
###and there're unlikely people underweight have hypertension 

#Relationship between gender & hypertension
compare_2 <- table(new_hyper$gender,new_hyper$hypertension)
mosaicplot(compare_2,color = c("light blue","purple"),
           main = "Hypertension vs gender")
###gender doesn't play an important role in hypertension
##which means both male and female have equal chances getting hypertension

####Relationship between gender & bmi
compare_3 <- table(new_hyper$bmi,new_hyper$gender)
mosaicplot(compare_3,color = c("light blue","light yellow"),
           main = "Bmi vs gender")
##overweight in male is more likely happend than in female

```

### *Hypertension vs bmi*

High percentage of people with hypertension without obesity, and there're unlikely people underweight have hypertension 

### *gender & hypertension*

gender doesn't play an important role in hypertension which means both male and female have equal chances getting hypertension

### *gender & bmi*

overweight in male is more likely happend than in female


